### 2021-04-22 report/submission.tex
Part 2 done.

			\begin{subfigure} {.475\columnwidth}
				\includegraphics[width=\columnwidth]{../plots/1ac_hist_randomwalk.png}
				\caption{Number distribution with distance of \(5 \times 10^5\) random walk particles}
				\label{fig:1c}
			\end{subfigure}%
			\hfill
			\begin{subfigure} {.475\columnwidth}
				\includegraphics[width=\columnwidth]{../plots/1ad_hist_randomwalk.png}
				\caption{Theoretical estimates and observed distribution of \(10^5\) random walk points. The blue histogram represents the observed number distribution , while the red line is a plot of analytical estimates.}
				\label{fig:1d}
			\end{subfigure}
			\caption{Problem 1.1}
			\label{fig:1}
		\end{figure}

@@

		\begin{figure} [h]
			\begin{subfigure} {.475\columnwidth}
				\includegraphics[width=\columnwidth]{../plots/2c_orbits.png}
				\caption{2c}
				\label{fig:2c}
			\end{subfigure}
			\hfill
			\begin{subfigure} {.475\columnwidth}
				\includegraphics[width=\columnwidth]{../plots/2d_energies.png}
				\caption{2d}
				\label{fig:2d}
			\end{subfigure}
			\caption{Problem 2}
			\label{fig:2}
		\end{figure}

@@

@@

@@

@@

@@

		\lstinputlisting[caption=first code, label=code:2.1,
						firstline=7,firstnumber=9,
						lastline=19]{../problem2/binary.m}
		\lstinputlisting[caption=second code, label=code2.2,
						firstline=4]{../problem2/energyplot.m}

@@

			\begin{table}
				\centering
				\begin{tabular} {l l r r}
					\toprule
					\textbf{Quantity} & \textbf{Variable} & \textbf{D=3} & \textbf{D=0.1}\\
					\midrule
					\(t^{}_{0}\) & \texttt{t} & -19.9555 & -15.1290 \\
					\(x^{}_{p}\) & \texttt{x(1)}& -400.0000 & -980.0000 \\
					\(y^{}_{p}\) & \texttt{x(2)} & -917.3151 & -199.7975 \\
					\(v^{}_{px}\) & \texttt{x(3)} & 37.6166 & 44.6972 \\
					\(v^{}_{py}\) & \texttt{x(4)} & 24.4949 & 4.4721 \\
					\(x^{}_{s}\) & \texttt{x(5)} & -400.0000 & -980.0000 \\
					\(y^{}_{s}\) & \texttt{x(6)} & -916.3151 & -198.7975 \\
					\(v^{}_{sx}\) & \texttt{x(7)} & 36.6166 & 43.6972 \\
					\(v^{}_{sy}\) & \texttt{x(8)} & 24.4949 & 4.4721 \\
					\bottomrule
				\end{tabular}
				\caption{Table of variable values for two values of D}
				\label{table:dvalues}
			\end{table}

>>>>

			\begin{subfigure} {.475\columnwidth}
				\includegraphics[width=\columnwidth]{../plots/1ac_hist_randomwalk.png}
				\caption{Number distribution with distance of \(5 \times 10^5\) random walk particles}
				\label{fig:1c}
			\end{subfigure}%
			\hfill
			\begin{subfigure} {.475\columnwidth}
				\includegraphics[width=\columnwidth]{../plots/1ad_hist_randomwalk.png}
				\caption{Theoretical estimates and observed distribution of \(10^5\) random walk points. The blue histogram represents the observed number distribution , while the red line is a plot of analytical estimates.}
				\label{fig:1d}
			\end{subfigure}
			\caption{Plotted results for problem 1.1A.}
			\label{fig:1}
		\end{figure}

@@

@@

			\item
				\subitem \textbf{Methods  --}
				The dimensionless masses and positions of primary and secondary stars are respectively given by,
				\begin{equation}
					\begin{gathered}
						\tilde{m}_p^{}=\frac{m_p^{}}{m} \text{ and } \tilde{m}_s^{}=\frac{m_s^{}}{m}, \\
						\tilde{x}_p^{}=\frac{x_p^{}}{a}, \text{ and } \tilde{x}_s^{}=\frac{x_s^{}}{a},
					\end{gathered}
					\label{eq:dimensionless_mass_pos}
				\end{equation}
				where, \(m\) and \(a\) are the total mass and separation of binary respectively. For a circular binary, the gravitational force experienced by each star equals the centripetal force on the star. Mathematically,
				\begin{equation}
					\begin{gathered}
						\frac{Gm_{s}^{}m_{p}^{}}{a^2_{}} = m_p^{} x_p^{} \omega^2_{} = m_s^{} x_s^{} \omega^2_{} , \\
						\Rightarrow x_s^{} = \frac{x_p^{} m_p^{}}{m_s^{}}.
					\end{gathered}
					\label{eq:secondary_pos}
				\end{equation}
				Total separation of binary is the sum of positions of primary and secondary stars with respect to their centre of mass, \textit{i.e.} 
				\begin{equation}
					x_p^{}+x_s^{} = a.
					\label{eq:binary_separation}
				\end{equation}
				Using equations \ref{eq:secondary_pos} and \ref{eq:binary_separation}, we obtain,
				\begin{equation}
					\begin{gathered}
						x_p + \frac{x_p m_p}{m_s} = a \\
						\Rightarrow x_p \left( \frac{m}{m_s}\right) = a \\
						\Rightarrow \tilde{x}_p = \tilde{m}_s 
					\end{gathered}
					\label{eq:xp_tilde}
				\end{equation}
				This relation gives us the relation for dimensionless position of primary star. Similarly, we can also obtain the relation for dimensionless position of secondary star as,
				\begin{equation}
					\begin{gathered}
						x_s + \frac{x_s m_s}{m_p} = a \\
						\Rightarrow \tilde{x}_s = \tilde{m}_p
					\end{gathered}
					\label{eq:xs_tilde}
				\end{equation}

				If we substitute the value of angular velocity, \(\omega =  v_p/r_p = v_s/r_s\), in equation \ref{eq:secondary_pos} for the two stars, we get,
				\begin{equation}
					\begin{gathered}
						\frac{G m_s m_p}{a^2} = \frac{m_p v_p^2}{x_p} = \frac{m_s v_s^2}{x_s} \\
						\Rightarrow \frac{v_p}{v_s} = \frac{m_s}{m_p}
					\end{gathered}
					\label{eq:gravity_centripetal}
				\end{equation}
				Multiplying both numerator and denominator of RHS by \(1/m\), we obtain,
				\begin{equation}
					\frac{v_p}{v_s} = \frac{\tilde{m}_s}{\tilde{m}_p}
					\label{eq:velocity_mass_relation}
				\end{equation}

				If we use the first relation of equation \ref{eq:gravity_centripetal}, we get,
				\begin{equation}
					\begin{gathered}
						\frac{G m_s m_p}{a^2} = \frac{m_p v_p^2}{x_p} \\
						\Rightarrow v_p^2 = \frac{G m_s x_p}{a^2} \\
						\Rightarrow v_p^2 = \frac{Gm}{a} \tilde{m}_s \tilde{x}_p = \frac{Gm}{a} \tilde{m}_s^2,
					\end{gathered}
				\end{equation}
				where, we have substituted values from equation set \ref{eq:dimensionless_mass_pos} and equation \ref{eq:xp_tilde}. In units of \(\sqrt{Gm/a}\), we obtain the value of dimensionless velocity of primary star as,
					\begin{equation}
						\tilde{v}_p = \tilde{m}_s
						\label{eq:primary_velocity}
					\end{equation}
				By using equations \ref{eq:gravity_centripetal} and \ref{eq:primary_velocity}, we get dimensionless velocity of secondary as,
					\begin{equation}
						\tilde{v}_s = \tilde{m}_p
						\label{eq:secondary_velocity}
					\end{equation}

				Given that angular velocity \(\omega = 2\pi / T\), where \(T\) is the period of the circular binary, we can find the dimensionless time by using equation \ref{eq:secondary_pos}, as,
				\begin{equation}
					\begin{gathered}
						\nonumber \frac{G m_s m_p}{a^2} = m_p x_p \omega^2 \\
						\Rightarrow \frac{G m_s}{a^2} = \frac{4 \pi^2 x_p}{T^2} \\
					\end{gathered}
				\end{equation}
				Using equations \ref{eq:dimensionless_mass_pos} and \ref{eq:xp_tilde}, we get,
				\begin{equation}
					\begin{gathered}
						\frac{G m_s}{a^3} = \frac{4 \pi^2 m_s}{T^2 m} \\
						\Rightarrow T = 2 \pi \sqrt{\frac{a^3}{Gm}}.
					\end{gathered}
				\end{equation}
				In units of \(\sqrt{a^3/Gm}\), we therefore derive dimensionless time of a circular binary as,
				\begin{equation}
					\tilde{T} = 2 \pi
					\label{eq:dimensionless_time}
				\end{equation}

				\subitem \textbf{Results  --}
				In the given scenario, the stars lie in x-y plane, with their centre of mass at origin. Primary star has initial coordinates \((-x_{px},0)\) while the secondary stars lies initially at \((x_{py},0)\).\\
				The stars rotate counter-clockwise, which gives us the velocity of primary star, \(-v_{py}\) and the velocity of secondary star, \(v_{sy}\).
				Using equations \ref{eq:xp_tilde}, \ref{eq:xs_tilde}, \ref{eq:primary_velocity} and \ref{eq:secondary_velocity}, the dimensionless values of initial conditions of the stars are summarized in table \ref{table:2_initial_conditions}.\\ The value of dimensionless time period of binary is derived in equation \ref{eq:dimensionless_time} as \(2 \pi\).

				\begin{table}
					\centering
					\begin{tabular} {c c c c}
						\toprule
						\textbf{Parameter} & \textbf{Value} & \textbf{Parameter} & \textbf{Value} \\
						(Primary Star) & & (Secondary star) & \\
						\midrule
						\(\tilde{x}_{px}\) & \(-\tilde{m}_s\) & \(\tilde{x}_{sx}\) & \(\tilde{m}_p\) \\
						\(\tilde{x}_{py}\) & 0 & \(\tilde{x}_{sy}\) & 0\\
						\(\tilde{v}_{px}\) & 0 & \(\tilde{v}_{sx}\) & 0 \\
						\(\tilde{v}_{py}\) & \(-\tilde{m}_s\) & \(\tilde{v}_{sy}\) & \(\tilde{m}_p\) \\
						\bottomrule
					\end{tabular}
				\caption{A summary of initial conditions for the simulation. All quantities are mentioned in dimensionless units, for a binary in x-y plane rotating counter-clockwise, with centre of mass at origin and primary star on the negative x-axis and secondary star on positive x-axis initially.}
				\label{table:2_initial_conditions}
				\end{table}

				%\subitem \textbf{Discussions  --}

@@

			\item
				\subitem \textbf{Methods  --}
				Given that,
				\begin{equation}
					\frac{m_p}{m_s} = 4
					\label{eq:masses_ratio}
				\end{equation}
				Adding 1 on both side, we get,
				\begin{equation}
					\begin{gathered}
						\nonumber \frac{m_p}{m_s} + 1 = 5 \\
						\Rightarrow \frac{m_p+m_s}{m_s} = \frac{m}{m_s} = 5
					\end{gathered}
				\end{equation}
				\begin{equation}
					\Rightarrow \tilde{m}_s = \frac{m_s}{m} = \frac{1}{5} = 0.2
					\label{eq:secondary_dimensionless_mass}
				\end{equation}

				Using equations \ref{eq:dimensionless_mass_pos}, \ref{eq:masses_ratio} and \ref{eq:secondary_dimensionless_mass}, we can obtain the dimensionless mass of primary star as,
				\begin{equation}
					\tilde{m_p} = 4 \times \tilde{m_s} = 0.8
					\label{eq:primary_dimensionless_mass}
				\end{equation}

				\subitem \textbf{Results  --}
				Using relations mentioned in table \ref{table:2_initial_conditions}, the initial conditions for the circular binary are defined in Matlab file \textit{binary.m} a shown in lines 13--20 of code \ref{code:2.1}. Using equations \ref{eq:secondary_dimensionless_mass} and \ref{eq:primary_dimensionless_mass}, the values of dimensionless masses of two stars are assigned in lines 11 and 12 as \texttt{mp = 0.8} and \texttt{ms = 0.2}.\\				
				The value of \texttt{tmax} is set ot \(10\pi\). Since from equation \ref{eq:dimensionless_time}, it is known that the value of dimensionless time is \(2\pi\), this given value of \texttt{tmax} will allow simulation to sun for 5 orbits of the binary.

				\begin{figure} [h] 
					\lstinputlisting[caption=binary.m - Initial conditions of circularly rotating binary for a simulation of five orbital rotations., label=code:2.1,
					firstline=7,firstnumber=9,
					lastline=19]{../problem2/binary.m}
				\end{figure}

%				\subitem \textbf{Discussions  --}

			\begin{figure} [h]
				\begin{subfigure} {.475\columnwidth}
					\includegraphics[width=\columnwidth]{../plots/2c_orbits.png}
					\caption{Trajectories of primary and secondary stars in a circular binary system over the first five rotations.}
					\label{fig:2c}
				\end{subfigure}
				\hfill
				\begin{subfigure} {.485\columnwidth}
					\includegraphics[width=\columnwidth]{../plots/2d_energies.png}
					\caption{Analytically estimated and numerically calculated evolution of orbital energy over five rotations. The lines are seen to overlap, indicating good agreement of theoretical expectations with numerical results.}
					\label{fig:2d}
				\end{subfigure}
				\caption{Plotted results for problem 2.}
				\label{fig:2}
			\end{figure}

@@

			\item
				\subitem \textbf{Methods  --}
				Executing the Matlab code resulted in creation of file named \emph{out}. This file is then used to plot the orbits of the two stars using the pre-existing program \emph{orbitplot.m}.

				\subitem \textbf{Results  --}
				The obtained plot, which shows the trajectories of the two stars over 5 rotations is shown in figure \ref{fig:2c}.

				\subitem \textbf{Discussions  --}
				It can be observed that dimensionless orbital radii of primary and secondary stars are \(0.2\) and \(0.8\) respectively. The orbits are circular around the centre of mass lying at origin.

@@

			\item
				\subitem \textbf{Methods  --}
				The energy of binary star system is given by,
				\begin{equation}
					E = \frac{1}{2} m_p v_p^2 + \frac{1}{2} m_s v_s^2 - \frac{G m_p m_s}{r}
					\label{eq:binary_energy}
				\end{equation}

				From equation \ref{eq:gravity_centripetal}, we van obtain relations for velocities of primary and secondary stars at any positions in their orbits as,
				\begin{equation}
					\begin{gathered}
						\frac{G m_s m_p}{r^2} = \frac{m_p v_p^2}{x_p} = \frac{m_s v_s^2}{x_s} \\
						\Rightarrow v_p^2 = \frac{G m_s x_p}{r^2} \text{ and } v_s^2 = \frac{G m_p x_s}{r^2}
					\end{gathered}
				\end{equation}

				Substituting these values in equation \ref{eq:binary_energy} and using binary separation, \(r = a = x_p+x_s\) from equation \ref{eq:binary_separation}, we get
				\begin{equation}
					\begin{gathered}
						\begin{aligned}
							E &= \frac{G m_p m_s x_p}{2a^2} + \frac{G m_s m_p x_s}{2a^2} - \frac{G m_p m_s}{a} \\
							&= \frac{G m_p m_s}{2a} - \frac{G m_p m_s}{a}\\
							&= -\frac{G m_p m_s}{2a} 
						\end{aligned}
					\end{gathered}
				\end{equation}

				In units of \(Gm^2/a\), the dimensionless energy is given by,
				\begin{equation}
					\tilde{E} = \cfrac{-\cfrac{G m_p m_s}{2a}}{\cfrac{G m^2}{a}} = -\cfrac{\tilde{m}_p \tilde{m}_s}{2}.
					\label{eq:dimensionless_energy}
				\end{equation}

				\begin{figure}
					\lstinputlisting[caption=second code, label=code:2.2,
					firstline=4]{../problem2/energyplot.m}
				\end{figure}

				The values of \(m_p\), \(m_s\) and \emph{tmax} are assigned to variables \texttt{mp}, \texttt{ms} and \texttt{tmax} in lines 1-3 of code \ref{code:2.2}. The derived value of dimensionless energy is specified in line 7 of the code. The program in code \ref{code:2.2} plots the time evolution of numerically and analytically calculated energies of the binary.

				\subitem \textbf{Results  --}
				The plot generated using code \ref{code:2.2} is shown in figure \ref{fig:2d}. The analytically estimated value of energy of the binary is given by equation \ref{eq:dimensionless_energy}.

				\subitem \textbf{Discussions  --}
				From equation \ref{eq:dimensionless_energy} we can conclude that the dimensionless energy of binary star system is constant and less than zero. In this case, its value from equation \ref{eq:dimensionless_energy} turns out to be \(-0.08\). Figure \ref{fig:2d} shows that numerically obtained energy of the system is similar to analytically calculated energy, both showing no variation in value over time.

@@

@@

			\begin{table}
				\centering
				\begin{tabular} {l l r r}
					\toprule
					\textbf{Quantity} & \textbf{Variable} & \textbf{D=3} & \textbf{D=0.1}\\
					\midrule
					\(t^{}_{0}\) & \texttt{t} & -19.9555 & -15.1290 \\
					\(x^{}_{p}\) & \texttt{x(1)}& -400.0000 & -980.0000 \\
					\(y^{}_{p}\) & \texttt{x(2)} & -917.3151 & -199.7975 \\
					\(v^{}_{px}\) & \texttt{x(3)} & 37.6166 & 44.6972 \\
					\(v^{}_{py}\) & \texttt{x(4)} & 24.4949 & 4.4721 \\
					\(x^{}_{s}\) & \texttt{x(5)} & -400.0000 & -980.0000 \\
					\(y^{}_{s}\) & \texttt{x(6)} & -916.3151 & -198.7975 \\
					\(v^{}_{sx}\) & \texttt{x(7)} & 36.6166 & 43.6972 \\
					\(v^{}_{sy}\) & \texttt{x(8)} & 24.4949 & 4.4721 \\
					\bottomrule
				\end{tabular}
				\caption{Table of variable values for two values of D}
				\label{table:3_dvalues}
			\end{table}
